---
title: "R Notebook"
output: html_notebook
---

This is the initial analysis of the COVID-19 data.
First, load the data - initially from the REDCap, but then from a data file.
Then choose composite outcome:

```{r}
library(tidyverse)
data <- data %>%
  mutate(composite = case_when(
    out_status.factor == "Non-survivor" |
      out_itu.factor == "Intensive care" ~
      1,
    TRUE ~ 0
  )) %>% mutate(composite_inc_niv = case_when(
    out_status.factor == "Non-survivor" |
      out_itu.factor == "Intensive care" |
      out_niv.factor != "No" ~
      1,
    TRUE ~ 0))

se <- function(x, ...) sqrt(var(x, ...)/length(x)) #alculate SE
```

Now, to make the QoL data.
First, select the columns, then make numeric, then drop na, for the moment.

```{r}
qol_scores <- data %>%
  select(293:328) %>%
  mutate_all(as.numeric) %>%
  drop_na()

qol_outputs <- sf36(qol_scores)
qol_outputs



```

 More work to make the QoL data - setting up the data etc.

```{r}
q <- qol_outputs %>%
  summarise(
    PF_m = mean(PF),
    PF_sd = sd(PF),
    RP_m = mean(RP),
    RP_sd = sd(RP),
    BP_m = mean(BP),
    BP_sd = sd(BP),
    GH_m = mean(GH),
    GH_sd = sd(GH),
    VT_m = mean(VT),
    VT_sd = sd(VT),
    SF_m = mean(SF),
    SF_sd = sd(SF),
    RD_m = mean(RE),
    RD_sd = sd(RE),
    MH_m = mean(MH),
    MH_sd = sd(MH),
    PCS_m = mean(PCS),
    PCS_sd = sd(PCS),
    MCS_m = mean(MCS),
    MCS_sd = sd(MCS)
    
    ) 

qol_outputs %>%
 pivot_longer(1:10, names_to = "questions", values_to = "result") %>%
  ggplot(aes(x = result, y = questions)) +
  geom_boxplot() +
  stat_summary(fun.y=mean, geom="point", shape=23, size=4) +
  coord_flip()

qol_outputs %>%
 pivot_longer(1:10, names_to = "questions", values_to = "result")

qol_outputs %>%
 pivot_longer(1:8, names_to = "questions", values_to = "result") %>%
  ggplot(aes(x = result, y = questions)) +
  geom_bar(stat = "summary", fun.y = "mean") + 
   geom_errorbar(stat='summary', width=.2) +
  coord_flip() +
  labs(title = "SF-36 mean and SD for the 8 domains") +
  theme_bw()

```


Now this makes the pretty graph: It seems pretty clear that QoL is affected by severity of disease...

```{r}


composite_out <- data %>%
  select(293:328, composite_inc_niv) %>%
  mutate_all(as.numeric) %>%
  drop_na() %>%
  select(composite_inc_niv) %>%
  bind_cols(qol_outputs) %>% mutate(composite_inc_niv = case_when(
    composite_inc_niv  ==1 ~ "NIV/ITU"
      ,
    TRUE ~"Ward only"))

composite_out %>%
  pivot_longer(2:11, names_to = "questions", values_to = "result") %>%
  mutate(composite_inc_niv =as.factor(composite_inc_niv)) %>%
  ggplot(aes(x = result, y = questions, fill = composite_inc_niv)) +
  geom_bar(stat = "summary", fun.y = "mean", position = "dodge") + 
   geom_errorbar(stat='summary', width=.2) +
    facet_wrap(~ composite_inc_niv) +
  coord_flip() +
  labs(title = "SF-36 mean and SD for the 8 domains") +
  theme_bw()
  

composite_out %>%
  pivot_longer(2:9, names_to = "questions", values_to = "result") %>%
    mutate(composite_inc_niv =as.factor(composite_inc_niv)) %>%
  group_by(composite_inc_niv, questions) %>%
  summarise(
    mean = mean(result),
    sd = se(result)
  ) %>% 
  ungroup() %>%
  ggplot(aes(x = questions, y = mean, fill = composite_inc_niv)) +
  geom_bar( stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymax = mean + sd, ymin = mean - sd), position = position_dodge(0.9), width = 0.2) +
  xlab("Domains of the SF-36") +
  ylab("Score (mean, standard error)")+  
  scale_fill_hue(name="Severity of disease")  +# Legend label, use darker colors
  theme_bw()

   
```


Now let's generate main database


```{r}
withqol <- data %>%
  drop_na(293:328) %>%
  bind_cols(qol_outputs)

withqol %>%
  pivot_longer(346:355, names_to = "questions", values_to = "result") %>%
    mutate(composite_inc_niv =as.factor(composite_inc_niv)) %>%
  group_by(composite_inc_niv, questions) %>%
  summarise(
    mean = mean(result),
    sd = se(result)
  ) %>% 
  ungroup() %>%
  ggplot(aes(x = questions, y = mean, fill = composite_inc_niv)) +
  geom_bar( stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymax = mean + sd, ymin = mean - sd), position = position_dodge(0.9), width = 0.2) +
  xlab("Domains of the SF-36") +
  ylab("Score (mean, standard error)")+  
  scale_fill_hue(name="Severity of disease")  +# Legend label, use darker colors
  theme_bw()


withqol %>%
  pivot_longer(346:355, names_to = "questions", values_to = "result") %>%
    mutate(composite_inc_niv =as.factor(composite_inc_niv)) %>%
  group_by(sex.factor, questions) %>%
  summarise(
    mean = mean(result),
    sd = se(result)
  ) %>% 
  ungroup() %>%
  ggplot(aes(x = questions, y = mean, fill = sex.factor)) +
  geom_bar( stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymax = mean + sd, ymin = mean - sd), position = position_dodge(0.9), width = 0.2) +
  xlab("Domains of the SF-36") +
  ylab("Score (mean, standard error)")+  
  scale_fill_hue(name="Gender")  +# Legend label, use darker colors
  theme_bw()





```


Now, let's look at QoL and antibody index. No correlation between antibody response and QoL - why would there be?

```{r}



cor(withqol$sp_abbottindex_12, withqol$PCS, use = "complete.obs")
cor(withqol$sp_abbottindex_12, withqol$MCS, use = "complete.obs")

ggplot(withqol) +
  geom_point(aes(x = sp_abbottindex_12, y = MCS, color = "Mental composite score")) +
  geom_point(aes(x = sp_abbottindex_12, y = PCS, color = "Physical composite score")) 
  

```


What about QoL and spirometry results

```{r}



cor(as.numeric(withqol$fu12_fvc), withqol$PCS, use = "complete.obs")
cor(withqol$sp_abbottindex_12, withqol$MCS, use = "complete.obs")

withqol %>%
  mutate(fu12_fvc = as.numeric(fu12_fvc)) %>%
  ggplot() +
  geom_point(aes(x = fu12_fvc, y = MCS, color = "Mental composite score")) +
  geom_point(aes(x = fu12_fvc, y = PCS, color = "Physical composite score")) 

withqol %>%
  mutate(fu12_fvcpred = as.numeric(fu12_fvcpred)) %>%
  ggplot() +
  geom_point(aes(x = fu12_fvcpred, y = MCS, color = "Mental composite score")) +
  geom_point(aes(x = fu12_fvcpred, y = PCS, color = "Physical composite score")) 


withqol %>%
  mutate_all(as.numeric) %>%
  pivot_longer(141:147, names_to = "questions", values_to = "result") %>%
  group_by(fu12_cxr, questions) %>%
  summarise(
    mean = mean(result, na.rm = TRUE),
    sd = se(result, na.rm = TRUE)
  ) %>% 
  ungroup() %>%
  mutate(fu12_cxr = as.factor(fu12_cxr)) %>%
  ggplot(aes(x = questions, y = mean, fill = fu12_cxr)) +
  geom_bar( stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymax = mean + sd, ymin = mean - sd), position = position_dodge(0.9), width = 0.2) +
  xlab("Domains of the SF-36") +
  ylab("Score (mean, standard error)")+  
  scale_fill_hue(name="Gender")  +# Legend label, use darker colors
  theme_bw()





```